---
title: "chromReads()"
author: "Adrià Mitjavila Ventura"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{chromReads()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `chromReads()`

`chromReads()` draws an `ggplot2`-based horizontal barplot with the number and percentage of reads mapped to each chromosome. 

It takes the path to a BAM file as an input. This BAM file must have a valid index.

```{r include=FALSE}
chromReads(bamfile = "../testdata/sampleX.bam")
```

```{r}
chromReads(bamfile = "../testdata/sampleX.bam")
```

The bar shows the reads mapped into each chromosome and the number at the right is the percentage of the total reads (including the filtered chromosomes).

## Filter chromosomes

Using the parameter `chr.filt`, the undesired chromosomes can be filtered out. 

This parameter requires a character vector with part of the names of the chromosomes to be filtered. Let's remove the *chr2L* and *chr2R* chromosomes:

```{r}
chromReads(bamfile = "../testdata/sampleX.bam", chr.filt = "chr2")
```

Now try removing the *chrX* and *chrY* chromosomes:

```{r}
chromReads(bamfile = "../testdata/sampleX.bam", chr.filt = c("X", "Y"))
```

Note that the percentage of reads does not change, since it's computed using all the reads from all the chromosomes.

## Add titles, etc

Using the parameters `main`, `subtitle`, `xlab` and `ylab` you can customize the title, the subtitle the X-axis label and the Y-axis label, respectively. 

```{r}
chromReads(bamfile = "../testdata/sampleX.bam", main = "This is a title", 
           subtitle = "This is a subtitle", xlab = "This the label of the X axis",
           ylab = "This is the label of the Y axis")
```

Moreover, you can customize the size of the previous mentioned labels, as well as the percentage label:

```{r}
chromReads(bamfile = "../testdata/sampleX.bam", main = "This is a title", 
           subtitle = "This is a subtitle", xlab = "This the label of the X axis",
           ylab = "This is the label of the Y axis", 
           main.size = 16, sub.size = 10, axis.size = 14, percent.size = 9)
```

## Remove percentage labels

By setting `percent.label` to FALSE, you can remove the percentage labels.

```{r}
chromReads(bamfile = "../testdata/sampleX.bam", percent.label = F)
```

## Further customization

Since `chromReads` outputs a `ggplot2`-based barplot, you can further customize it with `scales` or `theme`, etc.
