---
title: "barDEGs()"
author: "Adrià Mitjavila Ventura"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{barDEGs()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `barDEGs()`

`barDEGs()` draws an `ggplot2`-based horizontal barplot with the nubmer of upregulated and downregulated genes coming out from a DESeq analysis.

## Required

As input, `barDEGs` takes a named list of data frames with the columns `log2FoldChange`, `padj` and `DEG`. The `DEG` column must have the values *Upregulated*, *Downregulated* and *NS*.

```{r}
# read the dataframes into a list
deg_list <- list.files("../testdata", "diff_exp", full.names = T, recursive = T) %>%
  purrr::set_names(paste("cond", 1:3, sep = "")) %>%
  purrr::map(~read.delim(.x))

deg_list[[1]] %>% head()
```

## Minimum  run

```{r include=FALSE}
barDEGs(deg_list = deg_list)
```

```{r include=T}
barDEGs(deg_list = deg_list)
```

## Change thresholds

```{r}
barDEGs(deg_list = deg_list, pval = 0.1, log2FC = 1)
```

## Customize plot

### Names of conditions

```{r}
barDEGs(deg_list = deg_list, deg_names = c("Condition 1", "Condition 2", "Condition 3"))
```

### Position of condition labels

```{r}
barDEGs(deg_list = deg_list, name_pos = "right")
```

```{r}
barDEGs(deg_list = deg_list, name_pos = "left")
```

```{r}
barDEGs(deg_list = deg_list, name_pos = "none")
```

### Labels in Y-axis

```{r}
barDEGs(deg_list = deg_list, name_pos = "none", yaxis = T)
```

### DEGs in X-axis

```{r}
barDEGs(deg_list = deg_list, xaxis = T)
```

### X axis limits

```{r}
barDEGs(deg_list = deg_list, xlim = c(-1000, 1000))
```





### Further costumization

Since `barDEGs()` outputs a `ggplot2`-based barplot, it can be further customized with `scales` or `theme`, etc.
