---
title: "barDEGs()"
author: "Adrià Mitjavila Ventura"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{barDEGs()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 6, fig.height = 3, 
                      error = FALSE, message = FALSE, warning = FALSE)
devtools::load_all("..")
library(dplyr)
library(purrr)
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `barDEGs()`

`barDEGs()` draws an `ggplot2`-based horizontal barplot with the nubmer of upregulated and downregulated genes coming out from a DESeq analysis.

## Required inputs

As input, `barDEGs` takes a named list of data frames with the columns `log2FoldChange`, `padj` and `DEG`. The `DEG` column must have the values *Upregulated*, *Downregulated* and *NS*.

```{r}
# read the dataframes into a list
deg_list <- list.files("../testdata", "diff_exp", full.names = T, recursive = T) %>%
  purrr::set_names(paste("cond", 1:3, sep = "")) %>%
  purrr::map(~read.delim(.x))

deg_list[[1]] %>% head()
```

## Minimum  run

```{r include=T}
barDEGs(deg_list = deg_list)
```

## Change thresholds

By default, `barDEGs()` uses a `pval = 0.05` and `log2FC = 1`. If you want to set another thresholds to define the differentially expressed genes, you can change these parameters.

```{r}
barDEGs(deg_list = deg_list, pval = 0.1, log2FC = 1)
```

## Avoid thresholds

If your data frames don't have a *log2FoldChange* or *padj* columns, or if you do not want to use those columns, you can turn both `log2FC` and `pval` to `NULL` and they won't be used to define the differentially expressed genes.

Have in mind that, if you set `log2FC` and `pval` to `NULL`, your data frames must have a *DEG* column with the values *NS*, *Downregulated* and *Upregulated*.

```{r include=T}
# define a new list with different DEG values
# also remove the columns log2FoldChange and padj
deg_list_changed = deg_list %>% purrr::map(~dplyr::mutate(.x, DEG = ifelse(padj > 0.001, "NS", DEG)) %>% dplyr::select(-log2FoldChange, -padj))
  
# run barDEGs with log2FC and pval equal to NULL
barDEGs(deg_list = deg_list_changed, log2FC = NULL, pval = NULL)
```

Note that the number of differentially expressed genes has been reduced because we reduced the p-value threshold.

Also note that the data frames that were used lack the previously required columns *log2FoldChange* and *padj*.

## Count total

`barDEGs()` has the possibility to count the number of regions/genes in each element of the list and writting them to the caption by setting `total = TRUE`. In case all elements in `deg_list` have the same number of genes, the caption will be *N = <number_of_regions>*, otherwise it will be *<contrast1> = <number_of_regions1>; <contrast2> = <number_of_regions2>; ...*.

```{r include=T}
barDEGs(deg_list = deg_list, total = T)
```

## Compute p-value

Since `v4.0.0` it is possible to do a proportion test (`prop.test()`) to compute a p-value and look whether the proportion of up and downregulated genes is significantly different from a 0.5. 

To compute the p-value from the `prop.test()`, it is necessary to set `prop_test = T`. 



If `prop_test` is set to `TRUE`, the name of the contrast will be written to the left of the bars and the p-value will be written to the right.

```{r}
plot <- barDEGs(deg_list = deg_list, prop_test = T, xlim = c(-430,670))
plot
```

Briefly, `barDEGs()` takes the number of up and down regulated genes, the total number of differentially expressed genes and the expected proportions, which are 0.5. Inside `barDEGs()`, the `prop.test()` function is the following one:

```{r}
# Example of prop.test with one contrast (e.g. cond1)
data <- plot$data %>% dplyr::filter(contrast == "cond1")

# number of up and downregulated genes
up   <- data %>% dplyr::filter(DEG == "Upregulated") %>% dplyr::pull(n) 
down <- data %>% dplyr::filter(DEG == "Downregulated") %>% dplyr::pull(n) 

# number of differentially expressed genes
de = up+down

# do prop.test()
prop_test = prop.test(x = c(up,down), n = c(de,de), p = c(0.5,0.5))

# p-value
prop_test$p.value
```

## Customize plot

### Names of conditions

```{r}
barDEGs(deg_list = deg_list, deg_names = c("Condition 1", "Condition 2", "Condition 3"))
```

### Position of condition labels

```{r}
barDEGs(deg_list = deg_list, name_pos = "right")
```

```{r}
barDEGs(deg_list = deg_list, name_pos = "left")
```

```{r}
barDEGs(deg_list = deg_list, name_pos = "none")
```

### Labels in Y-axis

```{r}
barDEGs(deg_list = deg_list, name_pos = "none", yaxis = T)
```

### DEGs in X-axis

```{r}
barDEGs(deg_list = deg_list, xaxis = T)
```

### X axis limits

```{r}
barDEGs(deg_list = deg_list, xlim = c(-1000, 1000))
```

### Titles

`barDEGs()` has the possibility to add title and subtitle directly without adding any other layer. To add them, use the arguments `title` and `subtitle`.

```{r}
barDEGs(deg_list = deg_list, title = "This is a title", subtitle = "this is a subtilte")
```


### Further costumization

Since `barDEGs()` outputs a `ggplot2`-based barplot, it can be further customized with `scales` or `theme`, etc.
