---
title: "volcanoPlot()"
author: "Adrià Mitjavila Ventura"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{volcanoPlot()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `volcanoPlot()`

`volcanoPlot()` draws an `ggplot2`-based volcanoPlot with the *Log2(fold change)* at the X axis and the *-Log10(adjusted p-value)* in the Y axis.

## Required

As input, `barDEGs` takes a data frame with the columns `Geneid`, `log2FoldChange`, `padj` and `DEG`. The `DEG` column must have the values *Upregulated*, *Downregulated* and *NS*.

```{r}
# read the dataframes into a list
deg_list <- list.files("../testdata", "diff_exp", full.names = T, recursive = T) %>%
  purrr::set_names(paste("cond", 1:3, sep = "")) %>%
  purrr::map(~read.delim(.x))

deg_list[[1]] %>% head()
```

## Minimum  run

```{r include=FALSE, message=FALSE, warning=FALSE}
volcanoPlot(df = deg_list[[1]])
```

```{r include=T}
volcanoPlot(df = deg_list[[1]])
```

## Change thresholds

```{r}
volcanoPlot(df = deg_list[[1]], log2FC = 4, pval = 0.001)
```

## Customize plot

### Change axis limits

```{r}
volcanoPlot(df = deg_list[[1]], xlim = c(-5,5), ylim = c(0,10))
```

### Draw grid lines

```{r}
volcanoPlot(df = deg_list[[1]], gridLines = T)
```

### Specify titles

```{r}
volcanoPlot(df = deg_list[[1]], main = "This is a title", sub = "This is a subtitle", 
            xlab = "This is the X-axis label", ylab = "This is the Y-axis label")
```

```{r}
volcanoPlot(df = deg_list[[1]], main = "This is a title", sub = "This is a subtitle", 
            xlab = "This is the X-axis label", ylab = "This is the Y-axis label", 
            mainSize = 19, subSize = 15, axisLabelSize = 14, 
            axisTextSize = 15)
```

### Change DEG numbers

**Change size of DEG number labels**

```{r}
volcanoPlot(df = deg_list[[1]], labelSize = 12)
```

**Change color of DEG number labels**

```{r}
volcanoPlot(df = deg_list[[1]], labelColor = c("gold4", "blue"))
```

```{r}
volcanoPlot(df = deg_list[[1]], labelColor = c("black"))
```

**Change position of DEG number labels**

```{r}
volcanoPlot(df = deg_list[[1]], labelPos = 30)
```

### Change point colors

```{r}
volcanoPlot(df = deg_list[[1]], pointColor = c("cornflowerblue", "gray", "gold3"))
```

### Show genes names

**5 most significant genes**

```{r}
volcanoPlot(df = deg_list[[1]], degsLabel = T)
```

**Change number of genes to show**

```{r}
volcanoPlot(df = deg_list[[1]], degsLabel = T, degsLabelNum = 20)
```

**Show wanted genes**

```{r}
genes_to_show <- c("Lef1", "Mtor", "Ctnna1")

volcanoPlot(df = deg_list[[1]], degsLabel = genes_to_show, degsLabelSize = 5)
```

**Change size of gene labels**

```{r}
volcanoPlot(df = deg_list[[1]], degsLabel = T, degsLabelSize = 10)
```

### Further costumization

Since `barDEGs()` outputs a `ggplot2`-based scatter plot, it can be further customized with `scales` or `theme`, etc.
