---
title: "DEcompare()"
author: "Adrià Mitjavila Ventura"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{DEcompare()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `DEcompare()`

`DEcompare()` draws an `ggplot2`-based scatter plot comparing the *log2FC* of 2 different differential expression contrasts contrasts.

## Required

As input, `DEcompare()` takes a list of 2 data frames with the columns `Geneid`, `padj`.

```{r}
# read the dataframes into a list
deg_list <- list.files("../testdata", "diff_exp", full.names = T, recursive = T) %>%
  purrr::set_names(paste("cond", 1:3, sep = "")) %>%
  purrr::map(~read.delim(.x))

deg_list[[1]] %>% head()
```

## Minimum run

```{r include=FALSE}
DEcompare(deg_list = deg_list[1:2])
```

```{r include=T}
DEcompare(deg_list = deg_list[1:2])
```

## Change thresholds

```{r}
DEcompare(deg_list = deg_list[1:2], threshold = 0.1)
```

## Customize plot

### Change axis limits

```{r}
DEcompare(deg_list = deg_list[1:2], xlim = c(-5,5), ylim = c(-5, 7))
```

### Specify titles

```{r}
DEcompare(deg_list = deg_list[1:2], main = "This is a title", subtitle = "This is a subtitle", 
            xlab = "This is the X-axis label", ylab = "This is the Y-axis label")
```

### Change corner colors

```{r}
DEcompare(deg_list = deg_list[1:2], color_corners = c("Gray50", "Gray70", "Gray90", "White"))
```

```{r}
DEcompare(deg_list = deg_list[1:2], color_corners = c("White"))
```

```{r}
DEcompare(deg_list = deg_list[1:2], alpha_corners = c(1, .1, .4, .7))
```

```{r}
DEcompare(deg_list = deg_list[1:2], alpha_corners = .2)
```

### Further costumization

Since `DEcompare()` outputs a `ggplot2`-based scatter plot, it can be further customized with `scales` or `theme`, etc.
