---
title: "barAnno()"
author: "Adrià Mitjavila Ventura"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{barAnno()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)

peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x))
barAnno(anno_list = peak_list)
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `barAnno()` {.tabset}

`barAnno()` draws an `ggplot2`-based barplot ...

## Required input

As input, `barAnno()` takes a named list of data frames with the column `annotation`.

First, peak coordinates have to be read and the coordinate columns must be named as *seqnames*, *start* and *end*. ^Note: Since I will work with lists, I will use the `purrr` package for most of the data transformations.^

```{r}
# read the peaks into a list
peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x)) %>%
  # the files have the already annotated peaks, but we
  # will use only the coordinates to make the annotation ourselves
  # the columns seqnames, start and end are required
  purrr::map(~dplyr::select(.x, seqnames, start, end))

peak_list[[1]][1:5,]
```

Then, can use `ChIPseeker::annotatePeak()` to annotate the peaks into genomic regions.

```{r results='hide'}
# Load annotations for mouse mm10 genome
library(TxDb.Mmusculus.UCSC.mm10.knownGene) 
library(org.Mm.eg.db)

# Annotate peaks into csAnno object
peak_list_csAnno <- peak_list %>%
  # convert to granges
  purrr::map(~plyranges::as_granges(.x)) %>%
  # annotate peaks
  purrr::map(~ChIPseeker::annotatePeak(peak = .x, tssRegion = c(-2500, 2500), 
                                       TxDb = TxDb.Mmusculus.UCSC.mm10.knownGene,
                                       annoDb = "org.Mm.eg.db", level = "transcript"))
```

Now that peaks are annotated in the different genomic regions and we have created a *csAnno* object for each set of peaks, we must retrieve the `anno` data frame inside each *csAnno* object. This is not really necessary, since the corresponding transformations are already done inside `barAnno()`, but it is recommendable in case we want to do more things with these data.

```{r}
peak_list <- peak_list_csAnno %>% purrr::map(~as_tibble(.x))
```

## Default run

We can run `barAnno()` using as input:

* The list of data frames with the coordinates of the peaks and the annotated regions

```{r include=T}
barAnno(anno_list = peak_list)
```

* The list of *csAnno* objects that output from `ChIPseeker::annotatePeak()`

```{r include=T}
barAnno(anno_list = peak_list_csAnno)
```

## Plot number of peaks

```{r}
barAnno(anno_list = peak_list, fill_position = F)
```

## Genomic annotations

```{r}
barAnno(anno_list = peak_list, anno_num = 2) # default
```

```{r}
barAnno(anno_list = peak_list, anno_num = 3)
```

```{r}
barAnno(anno_list = peak_list, anno_num = "all")
```

## Change names

```{r}
barAnno(anno_list = peak_list, anno_names = c("Condition X", "Condition Y"))
```


```{r}
barAnno(anno_list = peak_list, 
        anno_names = c("Condition X", "Condition Y"), 
        names_order = c("Condition Y", "Condition X"))
```

## Customize plot

### Change bar width

```{r include=T}
barAnno(anno_list = peak_list, width = 0.6) # default
```

```{r include=T}
barAnno(anno_list = peak_list, width = 0.9)
```

```{r include=T}
barAnno(anno_list = peak_list, width = 1)
```

### Divide pannels

```{r}
barAnno(anno_list = peak_list, anno_names = c("Cond1", "Cond1"), protein = c("protX", "protY"))
```

```{r}
barAnno(anno_list = peak_list, anno_names = c("Cond1", "Cond1"), 
        protein = c("protX", "protY"), protein_order = c("protY", "protX"))
```

```{r}
barAnno(anno_list = peak_list, anno_names = c("Cond1", "Cond2"), protein = c("protX", "protY"))
```

### Set labels

```{r}
barAnno(anno_list = peak_list, main = "This is a title", subtitle = "This is a subtitle",
        xlab = "This is the X-axis label", ylab = "This is the Y-axis label")
```

### Legend position

```{r}
barAnno(anno_list = peak_list, legend_position = "none")
```

```{r}
barAnno(anno_list = peak_list, legend_position = "left")
```

```{r}
barAnno(anno_list = peak_list, legend_position = "bottom")
```

```{r}
barAnno(anno_list = peak_list, legend_position = "top")
```

### Change X text angle

```{r}
barAnno(anno_list = peak_list, xangle = 60)
```

### Change color palette

If the length of `color_palette` argument in `barAnno()` is 1 (e.g. `Set2`, the default), the function uses a predefined palette passed through `scale_fill_brewer()`. Available palettes can be found here: <https://ggplot2.tidyverse.org/reference/scale_brewer.html>. For example:

```{r}
barAnno(anno_list = peak_list, color_palette = "Set1")
```

```{r}
barAnno(anno_list = peak_list, color_palette = "Pastel2")
```

```{r}
barAnno(anno_list = peak_list, color_palette = "Oranges")
```

Nevertheless, if `color_palette` is a character vector of length greater than 1 and has valid color names (e.g. `c("blue", "gold3")`), the function takes these colors and passes them through `scale_fill_manual()` to fill the bars. If the number of annotated regions is greater than the default of 2 (i.e. `anno_num = 3` or `anno_num = "all"`), the number of colors passed through `color_palette` must be the same length.

```{r}
barAnno(anno_list = peak_list, color_palette = c("blue", "gold3"), anno_num = 2)
```

```{r}
barAnno(anno_list = peak_list, color_palette = c("blue", "gold3", "pink"), anno_num = 3)
```

```{r}
barAnno(anno_list = peak_list, anno_num = "all", 
        color_palette = c("blue", "gold3", "pink", "darkgreen", "darkred", "orange", "purple"))
```

The colors can be also generated through other functions, such as `rainbow()` or `circlize::rand_col()`.

```{r}
barAnno(anno_list = peak_list, anno_num = "all", 
        color_palette = rainbow(n = 7))
```


### Further costumization

Since `barAnno()` outputs a `ggplot2`-based bar plot, it can be further customized with `scales` or `theme`, etc.
