---
title: "barAnno()"
author: "Adrià Mitjavila Ventura"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{barAnno()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)

peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x))
barAnno(anno_list = peak_list)
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `barAnno()`

`barAnno()` draws an `ggplot2`-based barplot ...

## Required input

As input, `barAnno()` takes a named list of data frames with the column `annotation`.

```{r}
# read the peak annotation into a list
peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x))

peak_list[[1]][1:5, 1:7]
```

## Default run

```{r include=T}
barAnno(anno_list = peak_list)
```

## Number of peaks

```{r}
barAnno(anno_list = peak_list, fill_position = F)
```

## Genomic annotations

```{r}
barAnno(anno_list = peak_list, anno_num = 3)
```

```{r}
barAnno(anno_list = peak_list, anno_num = "all")
```

### Change names

```{r}
barAnno(anno_list = peak_list, anno_names = c("Condition X", "Condition Y"))
```


```{r}
barAnno(anno_list = peak_list, 
        anno_names = c("Condition X", "Condition Y"), 
        names_order = c("Condition Y", "Condition X"))
```

## Customize plot

### Divide pannels

```{r}
barAnno(anno_list = peak_list, anno_names = c("Cond1", "Cond1"), protein = c("protX", "protY"))
```

```{r}
barAnno(anno_list = peak_list, anno_names = c("Cond1", "Cond1"), 
        protein = c("protX", "protY"), protein_order = c("protY", "protX"))
```

```{r}
barAnno(anno_list = peak_list, anno_names = c("Cond1", "Cond2"), protein = c("protX", "protY"))
```



### Set titles

```{r}
barAnno(anno_list = peak_list, main = "This is a title", subtitle = "This is a subtitle",
        xlab = "This is the X-axis label", ylab = "This is the Y-axis label")
```

### Legend position

```{r}
barAnno(anno_list = peak_list, legend_position = "none")
```

```{r}
barAnno(anno_list = peak_list, legend_position = "left")
```

```{r}
barAnno(anno_list = peak_list, legend_position = "bottom")
```

```{r}
barAnno(anno_list = peak_list, legend_position = "top")
```

### Change X text angle

```{r}
barAnno(anno_list = peak_list, xangle = 60)
```

### Change color palette

Available palettes can be found here: <https://ggplot2.tidyverse.org/reference/scale_brewer.html>

```{r}
barAnno(anno_list = peak_list, color_palette = "Set1")
```

```{r}
barAnno(anno_list = peak_list, color_palette = "Pastel2")
```

```{r}
barAnno(anno_list = peak_list, color_palette = "Oranges")
```


### Further costumization

Since `barAnno()` outputs a `ggplot2`-based bar plot, it can be further customized with `scales` or `theme`, etc.
