---
title: "upsetPeaks()"
author: "Adrià Mitjavila Ventura"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{upsetPeaks()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)

peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x))  
upsetPeaks(peak_list)
```

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `upsetPeaks()`

`upsetPeaks()` draws an UpSet plot with the intersections between different sets of peaks using the function `getVennCounts()` (which calls `ChIPpeakAnno::makeVennDiagram()`) the and the package `UpSetR`.

Look at the [`UpSetR` package documentation](https://cran.rstudio.com/web/packages/UpSetR/).

Look at the [`makeVennDiagram()` function documentation](https://www.rdocumentation.org/packages/ChIPpeakAnno/versions/3.6.5/topics/makeVennDiagram)

## Required input

As input, `upsetPeaks()` takes a named list of data frames with the columns `seqnames`, `start` and `end`.

```{r}
# read the peak annotation into a list
peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x))

peak_list[[1]][1:5, 1:7]
```

## `getVennCounts()`

`getVennCounts()` calls `ChIPpeakAnno::makeVennDiagram()`, retrieves the Venn counts (number of overlaps between different sets of peaks) and builds a matrix of the peaks present in each set.

```{r}
venn_counts <- getVennCounts(peaks = peak_list)
```

```{r}
venn_counts$vennCounts
```

```{r}
venn_counts$matrix[1:5,]
```

`upsetPeaks()` calls the function `getVennCounts()` and builds the UpSet plot of the peaks using the Venn counts.

## Default run

```{r}
upsetPeaks(peak_list)
```

## Customize plot

### Change condition names

```{r}
upsetPeaks(peak_list, conds = c("Condition 1", "Condition 2"))
```

```{r}
upsetPeaks(peak_list, conds = c("Condition 1", "Condition 2"), conds_order = c("Condition 2", "Condition 1"))
```

### Different order

```{r}
upsetPeaks(peak_list, order.by = "freq") # default
upsetPeaks(peak_list, order.by = "degree")
upsetPeaks(peak_list, order.by = c("freq", "degree"))
```

### Change labels

```{r}
upsetPeaks(peak_list, mainbar.y.label = "This is an Y label for the main barplot", 
           sets.x.label = "This is the X label for the set size")
```
