---
title: "ggUpsetPeaks()"
author: "Adrià Mitjavila Ventura"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{ggUpsetPeaks()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)

peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x)) 
ggUpsetPeaks(peak_list)
```


# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `ggUpsetPeaks()`

`ggUpsetPeaks()` draws an UpSet plot with the intersections between different sets of peaks using the function `getVennCounts()` (which calls `ChIPpeakAnno::makeVennDiagram()`), builds each part of the plot separatelly using `ggplot2` and puts all plots together using `patchwork`.

Look at the [`patchwork` package documentation](https://patchwork.data-imaginist.com/)

Look at the [`makeVennDiagram()` function documentation](https://www.rdocumentation.org/packages/ChIPpeakAnno/versions/3.6.5/topics/makeVennDiagram)

## Required input

As input, `ggUpsetPeaks()` takes a named list of data frames with the columns `seqnames`, `start` and `end`.

```{r}
# read the peak annotation into a list
peak_list <- list.files("../testdata", "peak", full.names = T, recursive = T) %>%
  purrr::set_names(c("PeakX", "PeakY")) %>%
  purrr::map(~read.delim(.x))

peak_list[[1]][1:5, 1:7]
```

## `getVennCounts()`

`getVennCounts()` calls `ChIPpeakAnno::makeVennDiagram()`, retrieves the Venn counts (number of overlaps between different sets of peaks) and builds a matrix of the peaks present in each set.

```{r eval=FALSE}
venn_counts <- getVennCounts(peak_list)
```

```{r eval=FALSE}
venn_counts$vennCounts
#      PeakX PeakY Counts
# [1,]     0     0      0
# [2,]     0     1     70
# [3,]     1     0    977
# [4,]     1     1     23
# attr(,"class")
# [1] "VennCounts"
```

```{r eval=FALSE}
venn_counts$matrix[1:5,]
# peak    PeakX   PeakY
# peak1   0       1
# peak2	  0	      1
# peak3	  0	      1
# peak4	  0	      1
# peak5	  0	      1
```

`ggUpsetPeaks()` calls the function `getVennCounts()` and builds the UpSet plot of the peaks using the Venn counts.


## Default run

```{r}
ggUpsetPeaks(peak_list)
```

## Customize plot

### Condition names

```{r}
ggUpsetPeaks(peak_list, conds = c("Condition_1", "Condition_2"))
```

```{r}
#ggUpsetPeaks(peak_list, conds = c("Condition_1", "Condition_2"), conds_order = c("Condition_2", "Condition_1"))
```

### Don't sort by frequency

```{r}
ggUpsetPeaks(peak_list, order_by_freq = F)
```

### Size label

```{r}
ggUpsetPeaks(peak_list, num_size = 10)
```

### Set titles

```{r}
ggUpsetPeaks(peak_list, title = "This is a title", subtitle = "This is a subtitle", 
             caption = "This is a caption", )
```


### Further costumization

Since `ggUpsetPeaks()` outputs a `ggplot2`- and `patchwork`-based UpSet plot, it can be further customized with `scales` or `theme` and using the `patchwork` syntax.
