---
title: "expressionHeatmap2()"
author: "Adrià Mitjavila Ventura"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{expressionHeatmap2()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4)
devtools::load_all("..")
library(dplyr)
library(purrr)
degs_all <- list.files("../testdata", "diff", full.names = T, recursive = T) %>%
  purrr::map(~read.delim(.x)) %>%
  purrr::set_names(c("Cond1", "Cond2", "Cond3"))
plotmics::expressionHeatmap2(degs_all)
```


# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `expressionHeatmap2()`

`expressionHeatmap2()` takes a list of data frame with expression data for many genes and samples and draws a `ggplot2`-based heatmap. It also allows to cluster both rows (genes) and columns (samples), as well as scaling (`scale()`) by rows and columns (or both).


## Required input

As input, `expressionHeatmap()` takes a named list of data frames with expression data (i.e. *log2FoldChange*). Each data frame in the list corresponds to a condition, the first column -named *Geneid*- must have the gene names and the column with the values to be plotted must be named *log2FoldChange* (it can be another data than *log2FoldChange*, but the name is maintained because it is easier to plot log2FoldChange values coming from `DESeq2`).

```{r}
# read the deg data frames into a list
degs_all <- list.files("../testdata", "diff", full.names = T, recursive = T) %>%
  purrr::map(~read.delim(.x)) %>%
  purrr::set_names(c("Cond1", "Cond2", "Cond3"))

degs_all %>% purrr::map(~head(.x))
```

## Default run

To select the genes, the names written in the `genes` arguments must be present in the *Geneid* column of the input data.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"))
```

## Clustering

Setting the arguments `clust_rows` and `clust_cols` to `TRUE`, hierarchical clustering can be performed on both, rows and columns. 

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"),
                   legend_scale = c(-3,3),
                   clust_rows = T, clust_cols = T)
```

### Other clustering methods

By default, the distance calculation method is *euclidean* and the clustering is *ward.D*: `hclust(dist(data, method = "euclidean"), method = "ward.D")`. If another method has to be used, the `dist_method` and `hclust_method` arguments should be set to the corresponding value, passed through the functions `dist()` and `hclust()`, respectively. 

Available values for `dist_method` can be found in the [`dist()` function documentation](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/dist) and available values for `hclust_method` can be found in the [`hclust()` function documnetation](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/hclust).

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"),
                   legend_scale = c(-3,3),
                   clust_rows = T, clust_cols = T, 
                   dist_method = "manhattan", hclust_method = "median")
```

### Show dendograms

`expressionHeatmap2()` allows to plot the dendograms for the rows and the columns by setting the arguments `show_dend_rows` and `show_dend_cols` to TRUE, respectively. 

The dendograms are drawn with the function `plotDendogram()`, which takes the output of the `hclust()` function inside `expressionHeatmap2()` as input. Then, using the `pathcwork` package, the dendograms are attached to the main heatmap. Note that if the row dendogram is plotted (`show_dend_rows = T`), the Y axis is moved to the right.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"),
                   legend_scale = c(-3,3),
                   clust_rows = T, clust_cols = T,
                   show_dend_rows = T, show_dend_cols = T)
```

The size of the dendograms can be changed by setting their proportion to the height of the heatmap -for the column dendogram- or to its width -for the row dendogram-. To do so, the arguments `dend_rows_prop` and `dend_cols_prop` must be set to a value between 0 and 1.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"),
                   legend_scale = c(-3,3),
                   clust_rows = T, clust_cols = T,
                   show_dend_rows = T, show_dend_cols = T, 
                   dend_rows_prop = 0.5, dend_cols_prop = 0.5)
```

## Scaling

`expressionHeatmap2()` allows the scaling of the data by rows and columns by calling the function `scale()`. To do so, the `scale` argument must be set to either `"rows"` -to scale by rows- or `"cols"` -to scale by columns-. It can also be both `c("rows", "cols")`.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), scale = "rows")
```

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), scale = "cols")
```

## Legend scale

To change the legend scale, the arguments `legend_scale`, `legend_breaks_num`, `legend_breaks_by` and `legend_midpoint` can be used. 

By default, `legend_scale` is set to `c(-1.5,1.5)`. To set a custom scale, the `legend_scale` argument must be set to a numerical vector of length 2 (e.g. `c(-2,5)`). In such case, the argument `legend_breaks_by` will be used to define the distance between the breaks in the legend and the `legend_midpoint` argument will be set to the midpoint of the legend (passed through `scale_fill_gradient2(midpoint = legend_midpoint)`). 

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), 
                   legend_scale = c(-3,3), legend_breaks_by = 1, legend_midpoint = 0)
```

Note that the `legend_midpoint` value is the point where the central color is put, not necessarily the center of the legend.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), 
                   legend_scale = c(-3,3), legend_breaks_by = 1, legend_midpoint = 1)
```

We can also set `legend_scale` to NULL. This will take the lowest and highest values in the input data and uses them as lower limit and higher limit of the scale. In this case, the legend midpoint is set to half the way between the lower and higher limits (`legend_midpoint` won't be used).

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), 
                  legend_scale = NULL)
```

Even if `legend_scale = NULL`, using the argument `legend_breaks_num` instead of `legend_breaks_by` we can put as many breaks as we want. However, the breaks are rounded to the nearest integer so if our data has low values, it may happen that only few of the breaks appear. In such case, I recommend setting the `legend_scale` argument to a more adequate scale.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), 
                   legend_scale = NULL, legend_breaks_num = 1)
```


## Customize plot

### Write the values

By default, `expressionHeatmap2()` writes the expression values in each cell. This function can be disabled by setting the `write_label` to FALSE.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), write_label = F, 
                   legend_scale = c(-3,3))
```

If what is needed is to change the size, or color of the written values, the arguments `label_size` and `label_color` should be changed to the desired value.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), 
                   legend_scale = c(-3,3), label_size = 2, label_color = "darkred")
```

Furthermore, the number of decimals to round the expression values to can be changed using `label_digits`.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   label_digits = 0)
```

### Heatmap size

By default, the heatmap size is set to 10x10 mm for each cell. To change it, the arguments `hm_height` and `hm_width` must be set to the desired valuees in mm.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   hm_height = 70, hm_width = 70)
```

### Colors

To change the colors of the heatmap, the argument `hm_colors` will be used. This argument accepts a character vector with 3 colors: the first color will be the lower limit color, the second will be the midpoint color and the third will be the higher limit color. 

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   hm_colors = c("blue", "gray", "red"))
```

Note that if `legend_midpoint` value is the midpoint where the second color will be placed.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   legend_midpoint = 1, legend_breaks_by = 40, 
                   hm_colors = c("blue", "gray", "red"))
```

### Title, subtitle and caption

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   title = "This is a title", subtitle = "This is a subtitle", 
                   caption = "This is a caption")
```

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   title = "This is a title", subtitle = "This is a subtitle", 
                   caption = "This is a caption", title_hjust = .5)
```

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   title = "This is a title", title_size = 20,
                   subtitle = "This is a subtitle", subtitle_size = 18,
                   caption = "This is a caption", caption_size = 15)
```


### Axis

#### Axis titles

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   xlab = "X axis title", ylab = "Y axis title")
```


```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   xlab = "X axis title", ylab = "Y axis title", axis_title_size = 17)
```


#### Axis text size

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   axis_text_size = 17)
```

### Legend

#### Legend height

By default, the legend height is set to the same height as the heatmap, but it can be changed using the argument `legend_height`.

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   legend_height = 20)
```

#### Legend title

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   legend_title = "Legend Title")
```

```{r}
expressionHeatmap2(degs_all, genes = c("Lef1", "Jun", "Car4", "Yap1"), legend_scale = c(-3,3),
                   legend_title = "Legend Title", legend_title_size = 18)
```

### Further costumization

Since `expressionHeatmap()` outputs a `ggplot2`-based heatmap, it can be further customized like any other `ggplot2`-based plot. 

If dendograms are plotted, the customization will be done as in the `patchwork` package.
