---
title: "chromRegions()"
author: "Adrià Mitjavila Ventura"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{chromRegions()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F, message = F,
                      fig.align = "center", fig.width = 4, fig.height = 4)

devtools::load_all("..")
library(dplyr)

chrom_sizes = "../testdata/mm10.chrom.sizes"
regions      = "../testdata/mm10.regions.tsv"
regions2     = "../testdata/mm10.regions2.tsv"
plotmics::chromRegions(chrom_sizes = chrom_sizes, regions = regions)
```


# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `chromRegions()`

`chromRegions()` takes a file with the sizes of the chromosomes and draws a `ggplot2`-based bar plot. Then takes regions in a BED-like format to draw them into each chromosome. The inputs `chrom.sizes` and `regions` can be supplied as path to the file where the information is stored or a data frame.

## Required input

```{r}
chrom_sizes = "../testdata/mm10.chrom.sizes"
chrom_sizes %>% read.delim(header = F) %>% head()
```

```{r}
regions = "../testdata/mm10.regions.tsv"
regions %>% read.delim(header = F) %>% head()
```

## Default run

The default run requires only the `chrom_sizes` and the `regions` arguments, either as a path to a file or a data frame.

```{r}
chromRegions(chrom_sizes = "../testdata/mm10.chrom.sizes", regions = "../testdata/mm10.regions.tsv")
```

```{r}
chrom_sizes = "../testdata/mm10.chrom.sizes"%>% read.delim(header = F)
regions = "../testdata/mm10.regions.tsv" %>% read.delim(header = F)

chromRegions(chrom_sizes = chrom_sizes, regions = regions)
```

## Second set of regions

A second set of regions can be included in the plot. In that case, the default coloring by strand will be removed and the coloring will be done based on the region set.

```{r}
chromRegions(chrom_sizes = "../testdata/mm10.chrom.sizes", 
             regions     = "../testdata/mm10.regions.tsv",
             regions2    = "../testdata/mm10.regions2.tsv")
```

## Chromosome order

```{r}
chromRegions(chrom_sizes = "../testdata/mm10.chrom.sizes", 
             regions     = "../testdata/mm10.regions.tsv",
             chr_order   = c(1,3,4,2,5,6,7,9,8,19,11,10,12,13,15,14,16,17,18,"X","Y","M", "MT") )
```

## Exclude chromosomes

Very often, the genome assemblies of a lot of species have chromosomes/scaffolds with strange names, which are not nice to plot. These can be excluded using the `chr_exlude` argument with a vector of regular expressions that match the chromosomes to exclude. By default `chr_exclude` removes the most usuall *strange* chromosomes, but if you want to remove more chromosomes or don't want to remove any, you can change the `chr_exclude` argument. 

An example that excludes all the chromosomes that contain a dot in the name:

```{r}
chromRegions(chrom_sizes = "../testdata/mm10.chrom.sizes", 
             regions     = "../testdata/mm10.regions.tsv",
             chr_exclude = c("\\."))
```

Here you have an example that does not remove any chromosome.

```{r}
chromRegions(chrom_sizes = "../testdata/mm10.chrom.sizes", 
             regions     = "../testdata/mm10.regions.tsv",
             chr_exclude = c(" "))
```
Note that, since we are converting the chromosome names to an ordered factor with `chr_order`, the chromosome names that do not appear in `chr_order` will be groupped and plotted into a `NA` category.

## Customization

### Flip axes

`chromRegions()` allows flipping the axes by setting the `coord_flip` argument to TRUE.

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions      = "../testdata/mm10.regions.tsv",
             coord_flip   = T)
```

### Draw the points

By default, `chromRegions()` draws a line/rectangle and a point in the middle of each region. To avoid drawing the points, the argument `draw_points` can be set to FALSE.

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions      = "../testdata/mm10.regions.tsv",
             draw_points  = F)
```

### Color

By default, the regions are colored by strand, if the strand column has "+" and "-" as values. This coloring is controlled with `strand_color`, which must be a character vector with two colors.

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions      = "../testdata/mm10.regions.tsv",
             strand_color = c("Black", "Yellow"))
```
If you don't want to color by strand, just turn `strand_color` to NULL. In this case, the color of the region will be defined by `regions_colors`, which must be a character vector with two colors. In case that only one region is supplied (`regions` argument), only the first color will be used, but if a second set of regions is suppplied (`regions2` argument), both values will be used.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             strand_color  = NULL,
             regions_color = c("Gold3", "Darkgreen"))
```

If we supply a second set of regions:

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             regions2      = "../testdata/mm10.regions2.tsv",
             strand_color  = NULL,
             regions_color = c("Gold3", "Darkgreen"))
```
Now, imagine that we have regions that do not have a defined strand (e.g. most ChIP-seq peaks). In this case, the `strand_color` is internally converted to NULL and the `regions_colors` argument is used to color the regions.

```{r}
# Read and format regions file
regions_no_strand <- read.delim("../testdata/mm10.regions.tsv", header = F) %>% dplyr::mutate(V6 = ".")

# Draw the plot
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = regions_no_strand,
             strand_color  = c("Black", "Red"),
             regions_color = c("Gold3", "Darkgreen"))
```

### Titles

Title and subtitle can be supplied through the arguments `title` and `subtitle`, respectively. By default, they are set to NULL, but can accept a character of length 1.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             title         = "This is a title", 
             subtitle      = "This is a subtitle")
```

Also, the labels of the axes can be set or removed through the `xlab` and `ylab` arguments. By default, the X axis label is set to "Chromosome" and the Y axis is set to "Position (bp)", but they can be removed setting the corresponding arguments to NULL or changed to any value.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             xlab          = NULL, 
             ylab          = NULL)
```
Finally, a caption can be included in the bottom-right corner by setting the `caption` argument. By default, `caption` is set to NULL and it can be set to TRUE or any character. If `caption` is set to a character, whatever is written will be placed in the bottom-right corner. Instead, if it is set to TRUE, what will be written will be the number of regions in the input region sets.

Here there is an example with any character:

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             caption       = "This is a caption")
```

On the other hand, if `caption` is set to TRUE, the caption will show the number of regions in the input regions set (`regions`).

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             caption       = TRUE)
```

And, if `caption` is set to true, when two `regions2` is also supplied, the number of regions in each region set will be written in the caption.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             regions2      = "../testdata/mm10.regions2.tsv",
             caption       = TRUE)
```

### Legend 

The position of the legend is, by default, the bottom of the plot. This can be changed by changing the `legend` argument to one of "bottom", "right", "top", "left" or "none" (no legend). The `legend` argument is passed through `ggpubr::theme_pubr()`.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             legend        = "right")
```
```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions       = "../testdata/mm10.regions.tsv",
             legend        = "none")
```

### Further costumization

Since `chromRegions()` outputs a `ggplot2`-based bar plot, it can be further customized like any other `ggplot2`-based plot.
