---
title: "chromRegions()"
author: "Adrià Mitjavila Ventura"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{chromRegions()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F, message = F,
                      fig.align = "center", fig.width = 6, fig.height = 4)

devtools::load_all("..")
library(dplyr)

chrom_sizes = "../testdata/mm10.chrom.sizes"
regions_list <- list("Regions1" = "../testdata/mm10.regions.tsv",
                     "Regions2" = "../testdata/mm10.regions2.tsv",
                     "Regions3" = "../testdata/mm10.regions3.tsv")


plotmics::chromRegions(chrom_sizes = chrom_sizes, regions_list = regions_list)
```

Last updated: `r Sys.Date()`

***

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `chromRegions()`

`chromRegions()` takes a file with the sizes of the chromosomes and draws a `ggplot2`-based bar plot. Then takes a list of regions in a BED-like format to draw them into each chromosome. The inputs `chrom.sizes` and `regions` can be supplied as a list of character with path to the file where the information is stored or a list of data frame.

## Required input

The `crhom_sizes` input is:

```{r}
chrom_sizes = "../testdata/mm10.chrom.sizes"
chrom_sizes %>% read.delim(header = F) %>% head()
```

The `regions_list` is a list of characters or dataframes. It can have as many elements as wanted and it has the following structure:

```{r}
regions_list <- list("Regions1" = "../testdata/mm10.regions.tsv",
                     "Regions2" = "../testdata/mm10.regions2.tsv",
                     "Regions3" = "../testdata/mm10.regions3.tsv")
regions_list[[1]] %>% read.delim(header = F) %>% head()
```

## Default run

The default run requires only the `chrom_sizes` and the `regions` arguments, either as a path to a file or a data frame.

```{r}
chromRegions(chrom_sizes = "../testdata/mm10.chrom.sizes", regions_list = regions_list)
```

```{r}
# Read the data
chrom_sizes = "../testdata/mm10.chrom.sizes"%>% read.delim(header = F)
regions = regions_list %>% purrr::map(~read.delim(.x, header = F))

chromRegions(chrom_sizes = chrom_sizes, regions_list = regions)
```


## Chromosome order

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions_list = regions_list,
             chr_order    = c(1,3,4,2,5,6,7,9,8,19,11,10,12,13,15,14,16,17,18,"X","Y","M", "MT") )
```

## Exclude chromosomes

Very often, the genome assemblies of a lot of species have chromosomes/scaffolds with strange names, which are not nice to plot. These can be excluded using the `chr_exlude` argument with a vector of regular expressions that match the chromosomes to exclude. By default `chr_exclude` removes the most usuall *strange* chromosomes, but if you want to remove more chromosomes or don't want to remove any, you can change the `chr_exclude` argument. 

An example that excludes all the chromosomes that contain a dot in the name:

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions_list = regions_list,
             chr_exclude  = c("\\."))
```

Here you have an example that does not remove any chromosome.

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions_list = regions_list,
             chr_exclude  = c(" "))
```
Note that, since we are converting the chromosome names to an ordered factor with `chr_order`, the chromosome names that do not appear in `chr_order` will be groupped and plotted into a `NA` category.

## Customization

### Flip axes

`chromRegions()` allows flipping the axes by setting the `coord_flip` argument to TRUE.

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions_list = regions_list,
             coord_flip   = T)
```

### Draw the points

By default, `chromRegions()` draws a line/rectangle and a point in the middle of each region. To avoid drawing the points, the argument `draw_points` can be set to FALSE.

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions_list = regions_list,
             draw_points  = F)
```

### Color

By default, the regions are colored by region (i.e. each element in `regions_list`). This can be controlled with the `colors` argument, which accepts a character vector with valid color names and the same length as `regions_list` 

```{r}
chromRegions(chrom_sizes  = "../testdata/mm10.chrom.sizes", 
             regions_list = regions_list,
             colors       = c("Black", "Yellow", "Green"))
```
If you want to color by strand, just turn `color_by` to `"strand"`.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             colors        = c("Darkgreen", "Darkred"),
             color_by      = "strand")
```

Now, imagine that we have regions that do not have a defined strand (e.g. most ChIP-seq peaks). In this case, the `color_by` is internally converted to `"region"` and the regions are colored by region set (i.e. elements in `regions_list`). Look at this example with only one region set whose strand values are converted to ".".

```{r}
# Read and format regions file to have strand as "."
regions_no_strand <- list("Regions1" = read.delim("../testdata/mm10.regions.tsv", header = F) %>% dplyr::mutate(V6 = "."))

# Draw the plot
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_no_strand,
             color_by      = "strand",
             colors        = c("Gold3", "Darkgreen"))
```

### Titles

Title and subtitle can be supplied through the arguments `title` and `subtitle`, respectively. By default, they are set to NULL, but can accept a character of length 1.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             title         = "This is a title", 
             subtitle      = "This is a subtitle")
```

Also, the labels of the axes can be set or removed through the `xlab` and `ylab` arguments. By default, the X axis label is set to "Chromosome" and the Y axis is set to "", but they can be removed setting the corresponding arguments to NULL or changed to any value.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             xlab          = NULL,
             ylab          = "Position in the chromosome")
```
Finally, a caption can be included in the bottom-right corner by setting the `caption` argument. By default, `caption` is set to NULL and it can be set to TRUE or any character. If `caption` is set to a character, whatever is written will be placed in the bottom-right corner. Instead, if it is set to TRUE, what will be written will be the number of regions in the input region sets.

Here there is an example with any character:

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             caption       = "This is a caption")
```

On the other hand, if `caption` is set to TRUE, the caption will show the number of regions in the input regions set (`regions`).

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             caption       = TRUE)
```

### Legend 

The position of the legend is, by default, the bottom of the plot. This can be changed by changing the `legend` argument to one of "bottom", "right", "top", "left" or "none" (no legend). The `legend` argument is passed through `ggpubr::theme_pubr()`.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             legend        = "right")
```
```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             legend        = "none")
```
### Y axis

By default, the Y axis is not plotted, which means that the `size_y_text` is set to `NULL`. However, it can be plotted by setting the `size_y_text` argument to a number which will be used as size of the text in the Y axis.

```{r}
chromRegions(chrom_sizes   = "../testdata/mm10.chrom.sizes", 
             regions_list  = regions_list,
             size_y_text   = 10)
```


### Further costumization

Since `chromRegions()` outputs a `ggplot2`-based bar plot, it can be further customized like any other `ggplot2`-based plot.
