---
title: "ggVennBed()"
author: "Adrià Mitjavila Ventura"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{ggVennPeaks()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 4, fig.height = 4, 
                      warning = FALSE, error = FALSE, message = FALSE)
devtools::load_all("..")
library(dplyr)
library(purrr)
library(patchwork)

peak_list <- list.files("../testdata", "regions", full.names = T, recursive = T) %>%
  purrr::discard(~stringr::str_detect(string = .x, pattern = "tsv")) %>%
  purrr::set_names(c("PeakX", "PeakY")) 
```

Last updated: `r Sys.Date()`

***

# Load `plotmics`

```{r results=FALSE}
# load plotmics
library(plotmics)
```

# Run `ggVennBed()`

`ggVennBed()` draws an Venn diagram plot with the intersections of two BED files (A and B). It counts the number of unique regions in A and B, as well as the regions from A intersecting with B and viceversa. It uses the function `ggvenn()` from the `ggvenn` package, and the function `bt.intersect()` from the `bedtoolsr` package.

Look at the [`ggvenn` package documentation](https://github.com/yanlinlin82/ggvenn).

Look at the [`bedtoolsr` package documentation](https://github.com/PhanstielLab/bedtoolsr).

## Required input

As input, `ggVennBed()` takes two data frames in BED format or the paths to two BED-like files.

```{r}
# Example input. 
regionsX_path = "../testdata/regionsX.bed"
regionsY_path = "../testdata/regionsY.bed"

regionsX = read.delim(regionsX_path, header = F)
regionsY = read.delim(regionsY_path, header = F)

head(regionsX)
```

## Default run

You can run `ggVennBed` using the paths to two BED files:

```{r fig.height=7, fig.width=7}
ggVennBed(a = regionsX_path,  b = regionsY_path)
```

Or use two different data frames in BED-like format:

```{r}
ggVennBed(a = regionsX,  b = regionsY)
```

## Specify strandedness

By default, `ggVennBed()` intersects the regions with the same strandedness (`bedtools intersect -s -u -a regionA -b regionB`). To change this behaviour, change `stranded` to `FALSE` or `NULL`.

```{r}
ggVennBed(a = regionsX,  b = regionsY, stranded = TRUE) + ggtitle("Intersection with the same strandednes") +
ggVennBed(a = regionsX,  b = regionsY, stranded = FALSE) + ggtitle("Intersection regardless the strand")
```

## Customize plot

```{r}
ggVennBed(a = regionsX,  b = regionsY, title = "This is a title", title_size = 15, title_face = "bold")

ggVennBed(a = regionsX,  b = regionsY, title = "This is a title", subtitle = "This is a subtitle")
ggVennBed(a = regionsX,  b = regionsY, title = "This is a title", subtitle = "This is a subtitle", subtitle_size = 15, subtitle_face = "bold")

ggVennBed(a = regionsX,  b = regionsY, color = c("red", "green"))
ggVennBed(a = regionsX,  b = regionsY, color = c("red", "green"), alpha = 0.2)

ggVennBed(a = regionsX,  b = regionsY, labsize = 10)

ggVennBed(a = regionsX,  b = regionsY, setnames = c("RegionX", "RegionsY"))
ggVennBed(a = regionsX,  b = regionsY, setnames = c("RegionX", "RegionsY"), namesize = 10)

```


### Further costumization

Since `ggVennBed()` outputs a `ggvenn`-based Venn diagram and `ggvenn` is based in `ggplot2`, it can be further customized with `scales`, `theme`, etc...
