---
title: "expressionCor()"
author: "Adrià Mitjavila Ventura"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    self_contained: yes
    mode: selfcontained
vignette: >
  %\VignetteIndexEntry{expressionCor()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE, 
                      fig.width = 6, fig.height = 5, fig.align = "center")
```

# Load packages

```{r set_up, results=FALSE}
library(tidyverse) 
library(ggpubr)
library(patchwork)

library(plotmics)
```

# Run `expressionCor()`

## Required input

The input for `expressionCor()` is a data frame with a *Geneid* column and several columns with expression values of each sample. It is required that the values in the *Geneid* column are unique.

```{r}
expr <- read.delim("../testdata/tpm.tsv")
expr[1:5,1:5]
```

## Minimum run

```{r}
expressionCor(df = expr)
```

## Groups and facets

`expressionCor()` allows to use groups, which then will be used to separate the correlation plots into facets. To enable this, set the parameter `by_groups` to `TRUE`. This option requires an extra column named *group* in which each row will have a determined value. 

```{r fig.width=9, fig.height=3}
expr2 <- expr %>% dplyr::mutate(group = c(rep(c("group1", "group2", "group3"), nrow(expr)/3), "group1", "group2"))
head(expr2)
```


```{r fig.width=14, fig.height=5}
expressionCor(df = expr2, samples = colnames(expr)[-1], by_groups = T)
```

## Options

In `expressionCor()`, the default is to compute the Pearson correlation. Nevertheless, the correlation method can be changed to Spearman or Kendall correlations by changing the `corr_method` argument to `"spearman"` or `"kendall"`.

```{r fig.width=9, fig.height=4}
expressionCor(df = expr, corr_method = "pearson") + labs(title = "Pearson correlation (default)") +
expressionCor(df = expr, corr_method = "spearman") + labs(title = "Spearman correlation")
```

Also, the genes and/or samples used to compute the correlation values can be subset using the `genes` and `samples` parameters.

```{r fig.width=9, fig.height=4}
expressionCor(df = expr, genes = c("WASH7P", "FAM138A", "AL627309.1")) + labs(title = "Subset of genes") +
expressionCor(df = expr, samples = c("cond3_rep2", "cond1_rep1", "cond2_rep1")) + labs(title = "Subset of samples")
```

## Customization

### Heatmap

`expressionCor()` allows an easy customization of the heatmap:

* Plot type: 

  - by default, `expressionCor()` plots only the lower part of the matrix (since it is symmetrical), including the diagonal. 
  - setting the parameter `plot_type` to `"upper"` (plots the upper part of the matrix) or `"full"` (plots the whole matrix), the plot type can be changed.

```{r fig.width=12, fig.height=4}
expressionCor(df = expr, plot_type = "lower") + labs(title = "Lower plot (default)") +
expressionCor(df = expr, plot_type = "upper") + labs(title = "Upper plot") +
expressionCor(df = expr, plot_type = "full") + labs(title = "Full plot")
```

* Diagonal: 

  - `expressionCor()` plots the diagonal by default.
  - by setting `plot_diagonal = F`, the diagonal is not drawn.

```{r}
expressionCor(df = expr, plot_diagonal = F)
```

* Correlation values:
  
  - `expressionCor()` allows changing the coefficients color with the parameter `coeffs_color`, which is black by default.
  - the correlation coefficients can be removed from the plot if `coeffs_color` is set to `NULL`.
  - the parameter `coeffs_size` allows the customization of the coefficients size, which is 4 by default.

```{r fig.height=5, fig.width=11}
expressionCor(df = expr, coeffs_color = "Black") +
expressionCor(df = expr, coeffs_color = "Red", coeffs_size = 6)
```

* Colors:

  - the colors of the plot can be changed by setting the parameter `plot_colors` with a character vector with 3 valid R colors.
  - the values from `plot_colors` will be passed to `scale_fill_gradient2()` to set the color of the limits and the midpoint.
  
```{r}
expressionCor(expr, plot_colors = c("blue", "yellow", "red"))
```


* Size: 
  
  - the size of the heatmap (both with and height) can be changed with the parameter `plot_size` which, by default is 85 (millimeters).
  - changing the plot size, will automatically change the legend size to adjust it to the height (if the legend is vertical) or width (if the legend is horizontal).

```{r}
expressionCor(df = expr, plot_size = 50)
```


* Title, subtitle, caption:

  - the parameters `plot_title`, `plot_subtitle` and `plot_caption` allow to specify the title, subtitle and caption of the plot, respectively.
  - also, the text of the title, subtitle, caption can be formatted with the parameters, `title_hjust`, `title_face`, `title_size`, `subtitle_face`, `subtitle_size`, `caption_size`.

```{r}
expressionCor(df = expr, plot_title = "Title", plot_subtitle = "Subtitle", plot_caption = "Caption")
```

* Axes: 
  
  - the text in the axes can be formatted with `axis_text_color` and `axis_text_size`

```{r}
expressionCor(df = expr,axis_text_color = "blue", axis_text_size = 7)
```

* Cell border:

  - the parameter `cell_border` allows to set the color of the cell border which, by default is gray. It can be set to a valid R color or to a character "NA", which removes the border.

```{r fig.height=5, fig.width=15}
expressionCor(df = expr) +
expressionCor(df = expr, cell_border = "Red") +
expressionCor(df = expr, cell_border = "NA")
```

### Legend

`expressionCor()` allows an easy customization of the legend:

* Legend limits: 

  - by setting the parameters `legend_limits` and `legend_breaks_by`, you can change the legend scale which, by default, goes from -1 to 1.
  - the legend midpoint is automatically set to the mean of minimum and maximum values.

```{r}
expressionCor(df = expr, legend_limits = c(0.9,1), legend_breaks_by = 0.05)
```

* Legend title: 

  - change the `legend_title` parameter which, by default, is set to the correlation method used (e.g. if the correlation is Pearson's, the title is *Pearson correlation*).

```{r}
expressionCor(df = expr, legend_title  = "Title of the legend")
```

* Position of the legend: 

  - the `legend_pos` parameter allows to position the legend to the left, right, top or bottom.

```{r fig.height=11, fig.width=11}
(expressionCor(df = expr, legend_pos = "left", legend_title = "Legend at the left") +
expressionCor(df = expr, legend_pos = "right", legend_title = "Legend at the right")) /
(expressionCor(df = expr, legend_pos = "top", legend_title = "Legend at the top") +
expressionCor(df = expr, legend_pos = "bottom", legend_title = "Legend at the bottom"))
```

* Legend size: by setting the parameter `legend_size` with a numerical vector of length 2, the size of the legend can be changed. 
  
  - If the legend is vertical (i.e., at the left or at the right), the first number is the width and the second is the height of the legend, in millimeter. 
  - If the legend is horizontal (i.e., at the top or at the bottom), the first element is the height and the second is the width, in millimiters. 
  - By default, the legend size equals to the heatmap size. 

```{r fig.width=11, fig.height=5}
expressionCor(df = expr, legend_size = c(5,50)) +
expressionCor(df = expr, legend_pos = "bottom", legend_size = c(5,50))
```







